---
title: "check_dimensions"
author: "Anna Chinni"
format: pdf
editor: visual
---
1. loose vs strict

theoretical vs. practical
broad vs. specific
assimilated knowledge vs. direct evidence


2. Intuition-based vs. evidence-based X
3. Prescriptive vs. suggestive X
4. Rigid vs. flexible X
5. Situationally specific vs. broadly applicable X
6. Learned from authority vs. learned by experience X
abstract vs. concrete X
authority-based vs. experience-based X


```{r}
#Housekeeping
rm(list = ls())
graphics.off()

#load libraries
library(tidyverse)
library(ggimage)
library(factoextra)
library(ggplot2)

```
# Check loss
```{r}
d4 <- read_csv("log_4.csv")
```

```{r}
names(d4) <- gsub("^X\\.|\\.$", "", names(d4))
```

```{r}
d4 <- d4 %>% 
  mutate(
    dim = 4
  )
```

```{r}
d3 <- read_csv("log_3.csv")
```

```{r}
names(d3) <- gsub("^X\\.|\\.$", "", names(d3))
```

```{r}
d3 <- d3 %>% 
  mutate(
    dim = 3
  )
```

```{r}
d2 <- read_csv("log_2.csv")
```

```{r}
names(d2) <- gsub("^X\\.|\\.$", "", names(d2))
```

```{r}
d2 <- d2 %>% 
  mutate(
    dim = 2
  )
```

```{r}
d1 <- read_csv("log_1.csv")
```

```{r}
names(d1) <- gsub("^X\\.|\\.$", "", names(d1))
```

```{r}
d1 <- d1 %>% 
  mutate(
    dim = 1
  )
```

```{r}
d <- rbind(d1, d2, d3, d4)
```

```{r}
#just want final step loss
d <- d %>% 
  filter(step == "502")
```

```{r}
p_loss <- ggplot(d, aes(x = dim, y = loss_test)) +
  geom_line() +
  geom_point(size = 2) +
  labs(
    title = "Embedding Dimension vs Test Loss",
    x = "Numbers of Dimensions",
    y = "Test loss"
  ) +
  theme_classic()

p_loss
```
```{r}
ggsave("../plots/dimensions_loss.png", plot = p_loss, width = 6, height = 5, dpi = 300)
```

#load in embeddings

```{r}
#load in embedding
em4 <- read_csv("../results/embedding_4d.csv")
em3 <- read_csv("../results/embedding_3d.csv")
em2 <- read_csv("../results/embedding_2d.csv")
em1 <- read_csv("../results/embedding_1d.csv")
```
```{r}
#create color code

colors <-  tibble(
  word = em3$word,
  color = c("black", "black", "black", "black", "#7C3478", "#43597E", "#5479B2", "#6756B2", "#635392", "#7064B4", "#924DB6", "#A74BB1", "#924DB6", "#C55D9F", "#71285B", "#814597", "#675698")
)
```

#load in raw data
```{r}
data <- read_csv('../data/trials.csv')
```
#check how well each model predicts judgements
```{r}
add_embedding_distances <- function(data, em, prefix) {

  # get dimension columns
  dim_cols <- grep("^dim_", names(em), value = TRUE)

  # rename embeddings for joins
  em_head  <- em %>% rename_with(~ paste0("h_", .), all_of(dim_cols))
  em_win   <- em %>% rename_with(~ paste0("w_", .), all_of(dim_cols))
  em_lose  <- em %>% rename_with(~ paste0("l_", .), all_of(dim_cols))

  # join coordinates
  df <- data %>%
    left_join(em_head, by = c("head" = "word")) %>%
    left_join(em_win,  by = c("winner" = "word")) %>%
    left_join(em_lose, by = c("loser" = "word"))

  # compute vectors of column names
  h_cols <- paste0("h_", dim_cols)
  w_cols <- paste0("w_", dim_cols)
  l_cols <- paste0("l_", dim_cols)

  # Euclidean distances
  df[[paste0(prefix, "_dist_head_winner")]] <-
    sqrt(rowSums((df[h_cols] - df[w_cols])^2))

  df[[paste0(prefix, "_dist_head_loser")]] <-
    sqrt(rowSums((df[h_cols] - df[l_cols])^2))

  df
}
```
```{r}
data1 <- add_embedding_distances(data, em1, "em1")

data2 <- add_embedding_distances(data, em2, "em2")

data3 <- add_embedding_distances(data, em3, "em3")

data4 <- add_embedding_distances(data, em4, "em4")
```

```{r}
data1 <- data1 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em1_dist_head_winner, em1_dist_head_loser)

data2 <- data2 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em2_dist_head_winner, em2_dist_head_loser)

data3 <- data3 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em3_dist_head_winner, em3_dist_head_loser)

data4 <- data4 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em4_dist_head_winner, em4_dist_head_loser)
```

```{r}
data_all <- data %>%
  left_join(data1, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data2, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data3, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data4, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser"))
```

```{r}
data_all <- data_all %>% 
  mutate(
    em1_acc = if_else(em1_dist_head_winner < em1_dist_head_loser, 1, 0),
    em2_acc = if_else(em2_dist_head_winner < em1_dist_head_loser, 1, 0),
    em3_acc = if_else(em3_dist_head_winner < em1_dist_head_loser, 1, 0),
    em4_acc = if_else(em4_dist_head_winner < em1_dist_head_loser, 1, 0)
  )
```

```{r}
prop_acc_all <- tibble(
  em1 = mean(data_all$em1_acc),
  em2 = mean(data_all$em2_acc),
  em3 = mean(data_all$em3_acc),
  em4 = mean(data_all$em4_acc)
)
```

```{r}
data_filtered <- data_all %>% 
  filter(type == "random")
```

```{r}
prop_acc_filtered <- tibble(
  em1 = mean(data_filtered$em1_acc),
  em2 = mean(data_filtered$em2_acc),
  em3 = mean(data_filtered$em3_acc),
  em4 = mean(data_filtered$em4_acc)
)
```


```{r}
val <- data %>% 
  filter(type == "validation")
```

```{r}
majority_votes <- val %>%
  group_by(head, left, right) %>%
  count(winner, name = "n") %>%
  arrange(head, left, right, desc(n)) %>%
  slice(1) %>%
  ungroup() %>%
  rename(majority = winner)
```

```{r}
val <- val %>% 
  left_join(majority_votes %>% select(head, left, right, majority), by = c("head", "left", "right"))
```

```{r}
val <- val %>% 
  mutate(
    maj_agree = if_else(winner == majority, 1, 0),
    prop_agree = mean(maj_agree)
  )
```

```{r}
val_sum <- val %>% 
  group_by(head, left, right) %>% 
  summarise(
    prop_agree = mean(maj_agree)
  ) %>% 
  ungroup()
```

```{r}
prop_acc_filtered <- prop_acc_filtered %>% 
  mutate(
    human = unique(val$prop_agree)
  )
```

```{r}
prop_long <- prop_acc_filtered %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "prop_agree"
  )
```

```{r}
models <- prop_long %>% filter(model != "human")

human  <- prop_long %>% filter(model == "human")
```
## plot
```{r}
p_agree <- ggplot(models,
       aes(x = model,
           y = prop_agree)) +
  geom_hline(yintercept = human$prop_agree, linetype = "dashed") +
  annotate(
  "text",
  x = Inf,
  y = human$prop_agree,
  label = "Human agreement",
  vjust = 1.5,
  hjust = 1.1
)+
  geom_col(width = 0.5, fill = "steelblue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+ 
  labs(
    title = "Model Predictions",
    x = "Model",
    y = "Proportion agree with human judgement"
  ) +
 theme_classic()

p_agree
```
```{r}
ggsave("../plots/embedding_pred.png", plot = p_agree, width = 6, height = 5, dpi = 300)
```

# 3 Dimension Embedding

```{r}
#remove labels, so just embedding matrix
em3_matrix <- em3 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result3 <- prcomp(em3_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores3 <- as.data.frame(pca_result3$x)
```

```{r}
# Add words back in
pca_scores3$word <- em3$word

```

```{r}
write.csv(pca_scores3, "pca_3_dim.csv", row.names = FALSE)
```

```{r}
# Eigenvalues

eigenvalues3 <- (pca_result3$sdev)^2

```

```{r}
pca_df3 <- data.frame(
  Principal_Component = 1:length(eigenvalues3),
  Eigenvalue = eigenvalues3
)

```

```{r}
#vector format

#PC1 v 2
pc12_3 <- pca_scores3[, c("word", "PC1", "PC2")]

pc12_vec3 <- list(
  word = pca_scores3$word,
  PC1 = pca_scores3$PC1,
  PC2 = pca_scores3$PC2
)

#PC1 v 3
pc13_3 <- pca_scores3[, c("word", "PC1", "PC3")]

pc13_vec_3 <- list(
  word = pca_scores3$word,
  PC1 = pca_scores3$PC1,
  PC3 = pca_scores3$PC3
)
```

```{r}
# Scree plot

ggplot(pca_df3,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point() +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance3 <- pca_result3$sdev^2 /
  sum(pca_result3$sdev^2)
cumulative_variance3 <- cumsum(explained_variance3)
explained_variance_df3 <- data.frame(
  Dimension = 1:length(explained_variance3),
  ExplainedVariance = explained_variance3,
  CumulativeVariance = cumulative_variance3
)

print(explained_variance_df3)

# Plot cumulative variance explained

p3_variance <- ggplot(explained_variance_df3,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             color = "blue") +
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p3_variance

```
```{r}
ggsave("../plots/dime3_pca_variance.png", plot = p3_variance, width = 6, height = 5, dpi = 300)
```

##Plot in PCA space

```{r}
pca_scores3 <- pca_scores3 %>%
  left_join(colors, by = "word")
```

```{r}
# PC1 vs PC2

p3_pc12 <- ggplot(pca_scores3,
       aes(x = PC1,
           y = PC2,
           label = word,
           color = color)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5) +
  scale_color_identity() +
  labs(
    title = "PCA Space: PC1 vs PC2",
    x = "PC1",
    y = "PC2"
  ) +
theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p3_pc12
```

```{r}
# PC1 vs PC3

p3_pc13 <- ggplot(pca_scores3,
       aes(x = PC1,
           y = PC3,
           label = word,
           color = color)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC1 vs PC3",
    x = "PC1",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p3_pc13
```
```{r}
# PC1 vs PC3

p3_pc23 <- ggplot(pca_scores3,
       aes(x = PC2,
           y = PC3,
           label = word,
           color = color)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC2 vs PC3",
    x = "PC2",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p3_pc23
```


```{r}
ggsave("../plots/dim3_pca_pc1_pc2.png", plot = p3_pc12, width = 10, height = 9, dpi = 300)

ggsave("../plots/dim3_pca_pc1_pc3.png", plot = p3_pc13, width = 10, height = 9, dpi = 300)

ggsave("../plots/dim3_pca_pc2_pc3.png", plot = p3_pc23, width = 10, height = 9, dpi = 300)

ggsave("../plots/dim3_pca_pc1_pc2.pdf", plot = p3_pc12, width = 10, height = 9)

ggsave("../plots/dim3_pca_pc1_pc3.pdf", plot = p3_pc13, width = 10, height = 9)

ggsave("../plots/dim3_pca_pc2_pc3.pdf", plot = p3_pc23, width = 10, height = 9)
```

# 2 Dimension Embedding

```{r}
#remove labels, so just embedding matrix

em2_matrix <- em2 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result2 <- prcomp(em2_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores2 <- as.data.frame(pca_result2$x)
```

```{r}
# Add words back in
pca_scores2$word <- em2$word

```

```{r}
# Eigenvalues

eigenvalues2 <- (pca_result2$sdev)^2

```

```{r}
pca_df2 <- data.frame(
  Principal_Component = 1:length(eigenvalues2),
  Eigenvalue = eigenvalues2
)

```

```{r}
# Scree plot

ggplot(pca_df2,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point() +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance2 <- pca_result2$sdev^2 /
  sum(pca_result2$sdev^2)
cumulative_variance2 <- cumsum(explained_variance2)
explained_variance_df2 <- data.frame(
  Dimension = 1:length(explained_variance2),
  ExplainedVariance = explained_variance2,
  CumulativeVariance = cumulative_variance2
)

print(explained_variance_df2)

# Plot cumulative variance explained

p2_variance <- ggplot(explained_variance_df2,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             color = "blue") +
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p2_variance

```
```{r}
ggsave("../plots/dime2_pca_variance.png", plot = p2_variance, width = 6, height = 5, dpi = 300)
```

##Plot in PCA space
```{r}
pca_scores2 <- pca_scores2 %>%
  left_join(colors, by = "word")
```

```{r}
# PC1 vs PC2

p2_pc12 <- ggplot(pca_scores2,
       aes(x = PC1,
           y = PC2,
           label = word,
           color = color)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5) +
  scale_color_identity() +
  labs(
    title = "PCA Space: PC1 vs PC2",
    x = "PC1",
    y = "PC2"
  ) +
theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p2_pc12
```


```{r}
ggsave("../plots/dim2_pca_pc1_pc2.png", plot = p2_pc12, width = 6, height = 5, dpi = 300)

ggsave("../plots/dim2_pca_pc1_pc2.pdf", plot = p2_pc12, width = 6, height = 5)

```

# 1 Dimension Embedding
```{r}
#remove labels, so just embedding matrix

em1_matrix <- em1 %>%
  select(starts_with("dim_"))
```

```{r}
em1_plot <- em1 %>% 
  mutate(y = 0)
```

```{r}
em1_plot <- em1_plot %>%
  mutate(y_jitter = jitter(y, amount = 0.05))
```

```{r}
em1_plot <- em1_plot %>% 
  left_join(colors, by = "word")
```

##Plot 
```{r}
p_d1 <- ggplot(em1_plot,
       aes(x = dim_0,
           y = y_jitter,
           label = word,
           color = color)) +
  #geom_jitter(height = 0.15, width = 0, size = 3) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 5) +
  scale_y_continuous(limits = c(-.25, .25), expand = c(0,0))+ 
  scale_color_identity() +
  labs(
    title = "1 Dimension Embedding",
    x = "",
    y = ""
  ) +
theme_classic() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )


p_d1
```

```{r}
ggsave("../plots/dim1_embedding.png", plot = p_d1, width = 14, height = 13, dpi = 300)

ggsave("../plots/dim1_embedding.pdf", plot = p_d1, width = 14, height = 13)
```

#Participnts and Timing

```{r}
subj <- read_csv("../data/participants.csv")
```
```{r}
timing <- subj
```

```{r}
# Convert to datetime and calculate duration

timing <- timing %>%
  mutate(
    start_time = ymd_hms(started_at),
    stop_time  = ymd_hms(completed_at),
    
    # duration in seconds
    duration_sec = as.numeric(difftime(completed_at, started_at, units = "secs")),
    
    # duration in minutes
    duration_min = as.numeric(difftime(completed_at, started_at, units = "mins")),
    

    # duration in hours
    duration_hr = as.numeric(difftime(completed_at, started_at, units = "hours"))
  )

print(timing)

```

```{r}
timing <- timing %>%
  mutate(
    duration = completed_at - started_at
  )

print(timing$duration)
```

```{r}
mean(as.numeric(timing$duration_min)[as.numeric(timing$duration_min) < 200],
     na.rm = TRUE)
```
  