---
title: "check_dimensions"
author: "Anna Chinni"
format: pdf
editor: visual
---
1. loose vs strict

theoretical vs. practical
broad vs. specific
assimilated knowledge vs. direct evidence


2. Intuition-based vs. evidence-based X
3. Prescriptive vs. suggestive X
4. Rigid vs. flexible X
5. Situationally specific vs. broadly applicable X
6. Learned from authority vs. learned by experience X
abstract vs. concrete X
authority-based vs. experience-based X


```{r}
#Housekeeping
rm(list = ls())
graphics.off()

#load libraries
library(tidyverse)
library(ggimage)
library(factoextra)
library(ggplot2)

```
# Check loss

```{r}
data_dir <- "../results/check/logs"

files <- list.files(data_dir, full.names = TRUE)

df <- lapply(files, function(f) {
  d <- read.csv(f)
  d$run <- str_extract(basename(f), "(?<=run)\\d+(?=\\.csv)") %>% 
    as.integer()
  d
}) %>% 
  bind_rows()
```

```{r}
df <- df %>% 
  select(-run_id)
```

```{r}
df <- df %>%
  group_by(dim, run) %>% 
  filter(pf_calls == max(pf_calls))
```

```{r}
df_sum <- df %>% 
  group_by(dim) %>% 
  summarize(
    avg_loss = mean(loss_test),
    sd_loss = sd(loss_test),
    n = n(),
    se_loss = sd_loss/sqrt(n),
    ci_lower = avg_loss - 1.96 * se_loss,
    ci_upper = avg_loss + 1.96 * se_loss
  ) %>% 
  ungroup()
```


```{r}
p_loss <- ggplot(df_sum, aes(x = dim, y = avg_loss)) +
  geom_errorbar(
    aes(ymin = ci_lower,
        ymax = ci_upper),
    height = 0.2,
    width = .1,
    color = "black",
    linewidth = 0.2 
  ) +
  geom_line() +
  geom_point(size = 2) +
  labs(
    title = "Embedding Dimension vs Test Loss",
    x = "Numbers of Dimensions",
    y = "Test loss"
  ) +
  theme_classic()

p_loss
```
```{r}
ggsave("../plots/dimensions_loss.png", plot = p_loss, width = 6, height = 5, dpi = 300)
```

#load in embeddings

```{r}
#load in embedding
em5 <- read_csv("../results/embedding_5d.csv")
em4 <- read_csv("../results/embedding_4d.csv")
em3 <- read_csv("../results/embedding_3d.csv")
em2 <- read_csv("../results/embedding_2d.csv")
em1 <- read_csv("../results/embedding_1d.csv")
```
```{r}
#create color code

colors <-  tibble(
  word = em3$word,
  color = c("black", "black", "black", "black", "#7C3478", "#43597E", "#5479B2", "#6756B2", "#635392", "#7064B4", "#924DB6", "#A74BB1", "#924DB6", "#C55D9F", "#71285B", "#814597", "#675698")
)
```

#load in raw data
```{r}
data <- read_csv('../data/trials.csv')
```
#check how well each model predicts judgements
```{r}
add_embedding_distances <- function(data, em, prefix) {

  # get dimension columns
  dim_cols <- grep("^dim_", names(em), value = TRUE)

  # rename embeddings for joins
  em_head  <- em %>% rename_with(~ paste0("h_", .), all_of(dim_cols))
  em_win   <- em %>% rename_with(~ paste0("w_", .), all_of(dim_cols))
  em_lose  <- em %>% rename_with(~ paste0("l_", .), all_of(dim_cols))

  # join coordinates
  df <- data %>%
    left_join(em_head, by = c("head" = "word")) %>%
    left_join(em_win,  by = c("winner" = "word")) %>%
    left_join(em_lose, by = c("loser" = "word"))

  # compute vectors of column names
  h_cols <- paste0("h_", dim_cols)
  w_cols <- paste0("w_", dim_cols)
  l_cols <- paste0("l_", dim_cols)

  # Euclidean distances
  df[[paste0(prefix, "_dist_head_winner")]] <-
    sqrt(rowSums((df[h_cols] - df[w_cols])^2))

  df[[paste0(prefix, "_dist_head_loser")]] <-
    sqrt(rowSums((df[h_cols] - df[l_cols])^2))

  df
}
```
```{r}
data1 <- add_embedding_distances(data, em1, "em1")

data2 <- add_embedding_distances(data, em2, "em2")

data3 <- add_embedding_distances(data, em3, "em3")

data4 <- add_embedding_distances(data, em4, "em4")

data5 <- add_embedding_distances(data, em5, "em5")
```

```{r}
data1 <- data1 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em1_dist_head_winner, em1_dist_head_loser)

data2 <- data2 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em2_dist_head_winner, em2_dist_head_loser)

data3 <- data3 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em3_dist_head_winner, em3_dist_head_loser)

data4 <- data4 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em4_dist_head_winner, em4_dist_head_loser)

data5 <- data5 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em5_dist_head_winner, em5_dist_head_loser)
```

```{r}
data_all <- data %>%
  left_join(data1, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data2, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data3, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data4, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data5, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser"))
```

```{r}
data_all <- data_all %>% 
  mutate(
    em1_acc = if_else(em1_dist_head_winner < em1_dist_head_loser, 1, 0),
    em2_acc = if_else(em2_dist_head_winner < em2_dist_head_loser, 1, 0),
    em3_acc = if_else(em3_dist_head_winner < em3_dist_head_loser, 1, 0),
    em4_acc = if_else(em4_dist_head_winner < em4_dist_head_loser, 1, 0),
    em5_acc = if_else(em5_dist_head_winner < em5_dist_head_loser, 1, 0)
  )
```

```{r}
prop_acc_all <- tibble(
  em1 = mean(data_all$em1_acc),
  em2 = mean(data_all$em2_acc),
  em3 = mean(data_all$em3_acc),
  em4 = mean(data_all$em4_acc),
  em5 = mean(data_all$em5_acc)
)
```

```{r}
data_filtered <- data_all %>% 
  filter(type == "random")
```

```{r}
prop_acc_filtered <- tibble(
  em1 = mean(data_filtered$em1_acc),
  em2 = mean(data_filtered$em2_acc),
  em3 = mean(data_filtered$em3_acc),
  em4 = mean(data_filtered$em4_acc),
  em5 = mean(data_filtered$em5_acc)
)
```

```{r}
data_val <- data_all %>% 
  filter(type == "validation")

prop_acc_val <- tibble(
  em1 = mean(data_val$em1_acc),
  em2 = mean(data_val$em2_acc),
  em3 = mean(data_val$em3_acc),
  em4 = mean(data_val$em4_acc),
  em5 = mean(data_val$em5_acc)
)
  
```

```{r}
val <- data %>% 
  filter(type == "validation")
```

```{r}
majority_votes <- val %>%
  group_by(head, left, right) %>%
  count(winner, name = "n") %>%
  arrange(head, left, right, desc(n)) %>%
  slice(1) %>%
  ungroup() %>%
  rename(majority = winner)
```

```{r}
val <- val %>% 
  left_join(majority_votes %>% select(head, left, right, majority), by = c("head", "left", "right"))
```

```{r}
val <- val %>% 
  mutate(
    maj_agree = if_else(winner == majority, 1, 0),
    prop_agree = mean(maj_agree)
  )
```

```{r}
val_sum <- val %>% 
  group_by(head, left, right) %>% 
  summarise(
    prop_agree = mean(maj_agree)
  ) %>% 
  ungroup()
```

```{r}
prop_acc_filtered <- prop_acc_filtered %>% 
  mutate(
    human = unique(val$prop_agree)
  )
```

```{r}
prop_long <- prop_acc_filtered %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "prop_agree"
  )
```

```{r}
models <- prop_long %>% filter(model != "human")

human  <- prop_long %>% filter(model == "human")
```
## plot
```{r}
p_agree <- ggplot(models,
       aes(x = model,
           y = prop_agree)) +
  geom_hline(yintercept = human$prop_agree, linetype = "dashed") +
  annotate(
  "text",
  x = Inf,
  y = human$prop_agree,
  label = "Human agreement",
  vjust = 1.5,
  hjust = 1.1
)+
  geom_col(width = 0.5, fill = "steelblue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+ 
  labs(
    title = "Model Predictions",
    x = "Model",
    y = "Proportion agree with human judgement"
  ) +
 theme_classic()

p_agree
```
```{r}
ggsave("../plots/embedding_pred.png", plot = p_agree, width = 6, height = 5, dpi = 300)
```

# 5 Dimension Embedding

```{r}
#remove labels, so just embedding matrix
em5_matrix <- em5 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result5 <- prcomp(em5_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores5 <- as.data.frame(pca_result5$x)
```

```{r}
# Add words back in
pca_scores5$word <- em5$word

```

```{r}
write.csv(pca_scores5, "pca_5_dim.csv", row.names = FALSE)
```

```{r}
# Eigenvalues

eigenvalues5 <- (pca_result5$sdev)^2

```

```{r}
pca_df5 <- data.frame(
  Principal_Component = 1:length(eigenvalues5),
  Eigenvalue = eigenvalues5
)

```


```{r}
# Scree plot

ggplot(pca_df5,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point() +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance5 <- pca_result5$sdev^2 /
  sum(pca_result5$sdev^2)
cumulative_variance5 <- cumsum(explained_variance5)
explained_variance_df5 <- data.frame(
  Dimension = 1:length(explained_variance5),
  ExplainedVariance = explained_variance5,
  CumulativeVariance = cumulative_variance5
)

print(explained_variance_df5)

# Plot cumulative variance explained

p5_variance <- ggplot(explained_variance_df5,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             color = "blue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p5_variance

```

# 4 Dimension Embedding

```{r}
#remove labels, so just embedding matrix
em4_matrix <- em4 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result4 <- prcomp(em4_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores4 <- as.data.frame(pca_result4$x)
```

```{r}
# Add words back in
pca_scores4$word <- em4$word

```


```{r}
write.csv(pca_scores4, "pca_4_dim.csv", row.names = FALSE)
```

```{r}
#switch signs of PC1 and PC3 so correlations from scales are positive
pca_scores4 <- pca_scores4 %>% 
  mutate(
    PC1 = PC1*(-1),
    PC3 = PC3*(-1)
  )

```

```{r}
# Eigenvalues

eigenvalues4 <- (pca_result4$sdev)^2

```

```{r}
pca_df4 <- data.frame(
  Principal_Component = 1:length(eigenvalues4),
  Eigenvalue = eigenvalues4
)

```


```{r}
# Scree plot

ggplot(pca_df4,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point() +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance4 <- pca_result4$sdev^2 /
  sum(pca_result4$sdev^2)
cumulative_variance4 <- cumsum(explained_variance4)
explained_variance_df4 <- data.frame(
  Dimension = 1:length(explained_variance4),
  ExplainedVariance = explained_variance4,
  CumulativeVariance = cumulative_variance4
)

print(explained_variance_df4)

# Plot cumulative variance explained

p4_variance <- ggplot(explained_variance_df4,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             color = "blue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_variance

```
```{r}
pca_scores4 <- pca_scores4 %>%
  left_join(colors, by = "word")
```

```{r}
# PC1 vs PC2

p4_pc12 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC2,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  # scale_x_continuous(
  # limits = c(min(pca_scores4$PC1), max(pca_scores4$PC1) + 0.3)
  # ) +
  scale_y_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  labs(
    title = "PCA Space: PC1 vs PC2",
    x = "PC1",
    y = "PC2"
  ) +
theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc12
```

```{r}
# PC1 vs PC3

p4_pc13 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC3,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_y_continuous(
  limits = c(min(pca_scores4$PC3), max(pca_scores4$PC3) +.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC1 vs PC3",
    x = "PC1",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc13
```

```{r}
# PC1 vs PC4

p4_pc14 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC1 vs PC4",
    x = "PC1",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc14
```

```{r}
# PC2 vs PC3

p4_pc23 <- ggplot(pca_scores4,
       aes(x = PC2,
           y = PC3,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_x_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC2 vs PC3",
    x = "PC2",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc23
```

```{r}
# PC2 vs PC4

p4_pc24 <- ggplot(pca_scores4,
       aes(x = PC2,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_x_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC2 vs PC4",
    x = "PC2",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc24
```

```{r}
# PC3 vs PC4

p4_pc34 <- ggplot(pca_scores4,
       aes(x = PC3,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC3 vs PC4",
    x = "PC3",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc34
```

```{r}
ggsave("../plots/dim4_pca_pc1_pc2.png", plot = p4_pc12, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc3.png", plot = p4_pc13, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc4.png", plot = p4_pc14, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc2_pc3.png", plot = p4_pc23, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc2_pc4.png", plot = p4_pc24, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc3_pc4.png", plot = p4_pc34, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc2.pdf", plot = p4_pc12, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc1_pc3.pdf", plot = p4_pc13, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc1_pc4.pdf", plot = p4_pc14, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc2_pc3.pdf", plot = p4_pc23, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc2_pc4.pdf", plot = p4_pc24, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc3_pc4.pdf", plot = p4_pc34, width = 3.5, height = 3.5)
```


#Participnts and Timing

```{r}
subj <- read_csv("../data/participants.csv")
```
```{r}
timing <- subj
```

```{r}
# Convert to datetime and calculate duration

timing <- timing %>%
  mutate(
    start_time = ymd_hms(started_at),
    stop_time  = ymd_hms(completed_at),
    
    # duration in seconds
    duration_sec = as.numeric(difftime(completed_at, started_at, units = "secs")),
    
    # duration in minutes
    duration_min = as.numeric(difftime(completed_at, started_at, units = "mins")),
    

    # duration in hours
    duration_hr = as.numeric(difftime(completed_at, started_at, units = "hours"))
  )

print(timing)

```

```{r}
timing <- timing %>%
  mutate(
    duration = completed_at - started_at
  )

print(timing$duration)
```

```{r}
mean(as.numeric(timing$duration_min)[as.numeric(timing$duration_min) < 200],
     na.rm = TRUE)
```
  