---
title: "Scales"
author: "Anna Chinni"
format: pdf
editor: visual
---

# Setup

```{r}
#Housekeeping
rm(list = ls())
graphics.off()

#load libraries
library(tidyverse)
library(ggplot2)
library(skimr)
library(reshape2)
library(GPArotation)
library(psy)
library(plotrix)
library(FactoMineR)
library(tidytext)
```

```{r}
#load in data
demo <- read_csv("../data/scales/participants.csv")
d <- read_csv("../data/scales/trials.csv")
pca4 <- read_csv("pca_4_dim.csv")

```
We collected some demographic data we did not end up using in our analyses, but which could be identifying due to the specific population we sampled from. Therefore we remove those columns from our participant data here (commented out):

```{r}
# demo <- demo %>% 
#   select(-demo_languages)
```

```{r}
#write_csv(demo, "../data/scales/participants.csv")
```

```{r}
cat("We have data from", nrow(demo), "participants\n")
```

```{r}
d <- d %>% 
  mutate(
    response_align = if_else(scale_reversed == TRUE, response*(-1), response)
  )
```

At this point scales are ordered as:

loose to strict theoretical to practical assimilated knowledge to direct evidence specific to broad

```{r}
#switch signs of PC1 and PC3 so correlations from scales are positive
pca4 <- pca4 %>% 
  mutate(
    PC1 = PC1*(-1),
    PC3 = PC3*(-1)
  )

```

```{r}
d <- d %>% 
  mutate(response_align_scaled = ((response_align + 200) * 1)/400)
```

```{r}
#reorder scale_id levels to align with variance explained in PCA
d <- d %>% 
  mutate(
    scale_id = factor(scale_id, levels = c("strictness", "assim_direct", "broadness", "theory_practical"))
  )
```

```{r}
scales <- tibble(
  scale_id = c("theory_practical", "broadness", "assim_direct", "strictness"),
  low = c("theoretical", "specific", "assimilated knowledge", "loose"),
  high = c("practical", "broad", "direct evidence", "strict")
)

scales <- scales %>% 
  mutate(
    scale_id = factor(scale_id, levels = c("strictness", "assim_direct", "broadness", "theory_practical"))
  )
```

```{r}
d_sum <- d %>% 
  group_by(scale_id, word) %>% 
  summarize(
    avg_rating = mean(response_align_scaled),
    sd_rating = sd(response_align_scaled),
    n = n(),
    se_rating = sd_rating/sqrt(n),
    ci_lower = avg_rating - 1.96 * se_rating,
    ci_upper = avg_rating + 1.96 * se_rating
  ) %>% 
  ungroup()
```

```{r}
d_sum <- d_sum %>% 
  left_join(scales, by = "scale_id")
```

```{r}
#reorder scale_id levels to align with variance explained in PCA
d_sum <- d_sum %>% 
  mutate(
    scale_id = factor(scale_id, levels = c("strictness", "assim_direct", "broadness", "theory_practical"))
  )
```

```{r}
d_plot <- d_sum %>%
  mutate(word = reorder_within(word, avg_rating, scale_id))
```

```{r}
d_plot <- d_plot %>% 
  mutate(
    word_label = str_extract(word, "^[^_]+")
  )
```

# Plot ratings

```{r}
# p_scales <- ggplot(d_plot, aes(y = avg_rating, x = word)) +
#   geom_hline(yintercept = 0.5, color = "grey80", linewidth = 0.5) +
#   geom_errorbar(
#     aes(ymin = avg_rating - se_rating,
#         ymax = avg_rating + se_rating),
#     height = 0.2,
#     width = .2,
#     color = "darkgrey"
#   ) +
#   geom_point(size = 2, color = "darkgrey") +
#   facet_wrap(~ scale_id, ncol = 1, scales = "free_x") +
#   scale_x_reordered() +
#   scale_y_continuous(limits = c(-0.05, 1.05), expand = c(0,0))+
#   scale_x_discrete(expand = expansion(mult = c(0.01, 0.01))) +
#   labs(x = NULL, y = NULL) +
#   geom_text(
#     aes(label = word_label),
#     #hjust = 1,
#     #vjust = .5,
#     nudge_y = 0.05,
#     size = 4,
#     show.legend = FALSE
#   ) +
#   geom_text(
#     data = scales,
#     aes(x = -Inf, y = -Inf, label = low),
#     inherit.aes = FALSE,
#     hjust = 0, vjust = -.5,
#     size = 6
#   ) +
#   geom_text(
#     data = scales,
#     aes(x = -Inf, y = .95, label = high),
#     inherit.aes = FALSE,
#     hjust = 0, vjust = -.5,
#     size = 6
#   ) +
#   coord_cartesian(clip = "off") +
#   theme_classic() +
#   theme(
#     plot.margin = margin(10, 30, 10, 30),
#     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
#     #axis.text.x = element_blank(),
#     axis.ticks.x = element_blank(),
#     #axis.line.x = element_blank(),
#     panel.grid.major.x = element_blank(),
#     strip.text = element_blank(),
#     #strip.text = element_text(face = "bold"),
#     panel.spacing = unit(4, "lines")
#   )
# 
# p_scales
```

```{r}
p_scales <- ggplot(d_plot, aes(y = word, x = avg_rating)) +
  geom_vline(xintercept = 0.5, color = "grey80", linewidth = 0.25) +
  geom_errorbar(
    aes(xmin = ci_lower,
        xmax = ci_upper),
    height = 0.2,
    width = .2,
    color = "black",
    linewidth = 0.2 
  ) +
  geom_point(size = .5, color = "black") +
  facet_wrap(~ scale_id, ncol = 2, scales = "free_y") +
  scale_y_reordered() +
  scale_x_continuous(limits = c(0, 1), expand = c(0,0))+
  #scale_x_discrete(expand = expansion(mult = c(0.01, 0.01))) +
  #scale_x_discrete(expand = c(1,0)) +
  labs(x = NULL, y = NULL) +
  geom_text(
    data = scales,
    aes(x = -Inf, y = -Inf, label = low),
    inherit.aes = FALSE,
    hjust = 0, vjust = -.5,
    size = 4
  ) +
  geom_text(
    data = scales,
    aes(x = -Inf, y = .95, label = high),
    inherit.aes = FALSE,
    hjust = 0, vjust = -.5,
    size = 4
  ) +
  coord_cartesian(clip = "off") +
  theme_classic() +
  theme(
    plot.margin = margin(10, 30, 10, 30),
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 7),
    axis.line = element_line(linewidth = 0.3),
    axis.ticks = element_line(linewidth = 0.3),
    #axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1),
    #axis.text.x = element_blank(),
    axis.ticks.y = element_blank(),
    #axis.line.x = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.text = element_blank(),
    #strip.text = element_text(face = "bold"),
    panel.spacing = unit(4, "lines")
  )

p_scales
```

```{r}
ggsave("../plots/scales.pdf", plot = p_scales, width = 6, height = 4.3)
```

```{r}
levels(d_plot$scale_id)
```

# Participant distributions

```{r}
ggplot(
  d,
  aes(
    x = response_align_scaled,
    y = word
  )
) +
  geom_jitter(
    width = 0,
    height = 0.15,
    alpha = 0.4,
    size = 1.5
  ) +
  facet_wrap(~ scale_id, scales = "free_y") +
  labs(
    x = "Aligned response (scaled)",
    y = NULL
  ) +
  theme_bw()
```

# 4 Dimensions

```{r}
dim4 <- d_sum %>% 
  left_join(pca4, by = "word" )
```

```{r}
dim4 <- dim4 %>% 
  mutate(
    scale_id = factor(scale_id, levels = c("theory_practical", "broadness", "assim_direct", "strictness"))
  )
```

```{r}
dim4_corr <- dim4 %>% 
  group_by(scale_id) %>%
  summarise(
    cor_PC1 = cor.test(avg_rating, PC1, method = "pearson")$estimate,
    p_PC1   = cor.test(avg_rating, PC1, method = "pearson")$p.value,
    cor_PC2 = cor.test(avg_rating, PC2, method = "pearson")$estimate,
    p_PC2   = cor.test(avg_rating, PC2, method = "pearson")$p.value,
    cor_PC3 = cor.test(avg_rating, PC3, method = "pearson")$estimate,
    p_PC3   = cor.test(avg_rating, PC3, method = "pearson")$p.value,
    cor_PC4 = cor.test(avg_rating, PC4, method = "pearson")$estimate,
    p_PC4   = cor.test(avg_rating, PC4, method = "pearson")$p.value
  ) %>% 
  ungroup()
```

```{r}
dim4_corr_long <- dim4_corr %>%
  select(scale_id, starts_with("cor_")) %>%
  pivot_longer(
    cols = starts_with("cor_"),
    names_to = "PC",
    values_to = "correlation"
  ) %>%
  mutate(
    PC = str_remove(PC, "cor_")
  )
```

```{r}
dim4_corr_long <- dim4_corr %>%
  pivot_longer(
    -scale_id,
    names_to = c(".value", "PC"),
    names_sep = "_"
  )
```

```{r}
dim4_corr_long <- dim4_corr_long %>%
  mutate(
    label = paste0(
      round(cor, 2),
      "\n(",
      round(p, 3),
      ")"
    )
  )
```

```{r}
dim4_corr_long <- dim4_corr_long %>%
  mutate(
    sig = case_when(
      p < .001 ~ "***",
      p < .01  ~ "**",
      p < .05  ~ "*",
      TRUE     ~ ""
    ),
    label2 = paste0(round(cor, 2), sig)
  )
```

```{r}
p4 <- ggplot(dim4_corr_long,
       aes(x = PC,
           y = scale_id,
           fill = cor)) +
  geom_tile(color = "white") +
  geom_text(aes(label = label2),
            size = 2,
            lineheight = 0.9) +
  scale_fill_gradient2(
    low = "#af8dc3",
    high = "#7fbf7b",
    mid = "#f7f7f7",
    midpoint = 0,
    limits = c(-1, 1),
    name = "r"
  ) +
  theme_minimal(base_size = 9) +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme(
    panel.grid = element_blank(),
    # panel.border = element_rect(
    #   color = "black",
    #   fill = NA,
    #   linewidth = 1
    # )
  )

p4
```

```{r}
ggsave("../plots/dim4_correlations.pdf", plot = p4, width = 3, height = 1)
```

```{r}
```
