---
title: "check_dimensions"
author: "Anna Chinni"
format: pdf
editor: visual
---

# Setup 
```{r}
#Housekeeping
rm(list = ls())
graphics.off()

#load libraries
library(tidyverse)
library(ggimage)
library(factoextra)
library(ggplot2)

```

We collected some demographic data we did not end up using in our analyses, but which could be identifying due to the specific population we sampled from. Therefore we remove those columns from our participant data here (commented out):

```{r}
demo <- read_csv("../data/participants.csv")
```

```{r}
# demo <- demo %>% 
#   select(-demo_languages, -demo_age)
```

```{r}
#write_csv(demo, "../data/participants.csv")
```

# Check loss
We check the average loss of solutions as each dimensionality (1, 2, 3, 4, and 5 dimensions). We do this by estimating each embedding 10 times and then recording the final loss of the models. Lower loss indicate lower error and better fit to the data. We calculate 95% confidence intervals for each model. 
```{r}
data_dir <- "../results/check/logs"

files <- list.files(data_dir, full.names = TRUE)

df <- lapply(files, function(f) {
  d <- read.csv(f)
  d$run <- str_extract(basename(f), "(?<=run)\\d+(?=\\.csv)") %>% 
    as.integer()
  d
}) %>% 
  bind_rows()
```

```{r}
df <- df %>% 
  select(-run_id)
```

```{r}
df <- df %>%
  group_by(dim, run) %>% 
  filter(pf_calls == max(pf_calls))
```

```{r}
df_sum <- df %>% 
  group_by(dim) %>% 
  summarize(
    avg_loss = mean(loss_test),
    sd_loss = sd(loss_test),
    n = n(),
    se_loss = sd_loss/sqrt(n),
    ci_lower = avg_loss - 1.96 * se_loss,
    ci_upper = avg_loss + 1.96 * se_loss
  ) %>% 
  ungroup()
```

```{r}
p_loss <- ggplot(df_sum, aes(x = dim, y = avg_loss)) +
  geom_errorbar(
    aes(ymin = ci_lower,
        ymax = ci_upper),
    height = 0.2,
    width = .1,
    color = "black",
    linewidth = 0.2 
  ) +
  geom_line() +
  geom_point(size = .5) +
  labs(
    title = "Embedding Dimension vs Test Loss",
    x = "Numbers of Dimensions",
    y = "Test loss"
  ) +
  theme_classic()

p_loss
```

```{r}
ggsave("../plots/dimensions_loss.pdf", plot = p_loss, width = 3, height = 2)
```

#load in embeddings

We estimated embedings at 5 dimentionalities (1, 2, 3, 4, and 5)
```{r}
#load in embedding - these are the first embeddings run for each dimensionality
em5 <- read_csv("../results/embedding_5d.csv")
em4 <- read_csv("../results/embedding_4d.csv")
em3 <- read_csv("../results/embedding_3d.csv")
em2 <- read_csv("../results/embedding_2d.csv")
em1 <- read_csv("../results/embedding_1d.csv")
```

```{r}
#create color code

colors <-  tibble(
  word = em3$word,
  color = c("black", "black", "black", "black", "#7C3478", "#43597E", "#5479B2", "#6756B2", "#635392", "#7064B4", "#924DB6", "#A74BB1", "#924DB6", "#C55D9F", "#71285B", "#814597", "#675698")
)
```

#load in raw data

```{r}
data <- read_csv('../data/trials.csv')
```

#check how well each model predicts judgements
We calculate the euclidean distance between each choice from the target per triplet. We define the model's "choice" on a given trial as the choice with the shorter euclidean distance from the target. We compare this against the actual choices participants made for eahc trial. 
```{r}
add_embedding_distances <- function(data, em, prefix) {

  # get dimension columns
  dim_cols <- grep("^dim_", names(em), value = TRUE)

  # rename embeddings for joins
  em_head  <- em %>% rename_with(~ paste0("h_", .), all_of(dim_cols))
  em_win   <- em %>% rename_with(~ paste0("w_", .), all_of(dim_cols))
  em_lose  <- em %>% rename_with(~ paste0("l_", .), all_of(dim_cols))

  # join coordinates
  df <- data %>%
    left_join(em_head, by = c("head" = "word")) %>%
    left_join(em_win,  by = c("winner" = "word")) %>%
    left_join(em_lose, by = c("loser" = "word"))

  # compute vectors of column names
  h_cols <- paste0("h_", dim_cols)
  w_cols <- paste0("w_", dim_cols)
  l_cols <- paste0("l_", dim_cols)

  # Euclidean distances
  df[[paste0(prefix, "_dist_head_winner")]] <-
    sqrt(rowSums((df[h_cols] - df[w_cols])^2))

  df[[paste0(prefix, "_dist_head_loser")]] <-
    sqrt(rowSums((df[h_cols] - df[l_cols])^2))

  df
}
```

```{r}
data1 <- add_embedding_distances(data, em1, "em1")

data2 <- add_embedding_distances(data, em2, "em2")

data3 <- add_embedding_distances(data, em3, "em3")

data4 <- add_embedding_distances(data, em4, "em4")

data5 <- add_embedding_distances(data, em5, "em5")
```

```{r}
data1 <- data1 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em1_dist_head_winner, em1_dist_head_loser)

data2 <- data2 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em2_dist_head_winner, em2_dist_head_loser)

data3 <- data3 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em3_dist_head_winner, em3_dist_head_loser)

data4 <- data4 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em4_dist_head_winner, em4_dist_head_loser)

data5 <- data5 %>% 
  select(participant_id, trial_index, head, left, right, winner, loser, em5_dist_head_winner, em5_dist_head_loser)
```

```{r}
data_all <- data %>%
  left_join(data1, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data2, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data3, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data4, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser")) %>%
  left_join(data5, by = c("participant_id", "trial_index", "head", "left", "right", "winner", "loser"))
```

```{r}
data_all <- data_all %>% 
  mutate(
    em1_acc = if_else(em1_dist_head_winner < em1_dist_head_loser, 1, 0),
    em2_acc = if_else(em2_dist_head_winner < em2_dist_head_loser, 1, 0),
    em3_acc = if_else(em3_dist_head_winner < em3_dist_head_loser, 1, 0),
    em4_acc = if_else(em4_dist_head_winner < em4_dist_head_loser, 1, 0),
    em5_acc = if_else(em5_dist_head_winner < em5_dist_head_loser, 1, 0)
  )
```

```{r}
prop_acc_all <- tibble(
  em1 = mean(data_all$em1_acc),
  em2 = mean(data_all$em2_acc),
  em3 = mean(data_all$em3_acc),
  em4 = mean(data_all$em4_acc),
  em5 = mean(data_all$em5_acc)
)
```

```{r}
prop_acc_all2 <- data_all %>% 
  summarize(
  n = n(),
  
  em1_prop = mean(em1_acc),
  em1_sd = sd(em1_acc),
  em1_se = em1_sd/sqrt(n),
  em1_ci_lower = em1_prop - 1.96 * em1_se,
  em1_ci_upper = em1_prop + 1.96 * em1_se,
  
  em2_prop = mean(em2_acc),
  em2_sd = sd(em2_acc),
  em2_se = em2_sd/sqrt(n),
  em2_ci_lower = em2_prop - 1.96 * em2_se,
  em2_ci_upper = em2_prop + 1.96 * em2_se,
  
  em3_prop = mean(em3_acc),
  em3_sd = sd(em3_acc),
  em3_se = em3_sd/sqrt(n),
  em3_ci_lower = em3_prop - 1.96 * em3_se,
  em3_ci_upper = em3_prop + 1.96 * em3_se,
  
  em4_prop = mean(em4_acc),
  em4_sd = sd(em4_acc),
  em4_se = em4_sd/sqrt(n),
  em4_ci_lower = em4_prop - 1.96 * em4_se,
  em4_ci_upper = em4_prop + 1.96 * em4_se,
  
  em5_prop = mean(em5_acc),
  em5_sd = sd(em5_acc),
  em5_se = em5_sd/sqrt(n),
  em5_ci_lower = em5_prop - 1.96 * em5_se,
  em5_ci_upper = em5_prop + 1.96 * em5_se
  )

```

```{r}
data_filtered <- data_all %>% 
  filter(type == "random")
```

```{r}
prop_acc_filtered <- tibble(
  em1 = mean(data_filtered$em1_acc),
  em2 = mean(data_filtered$em2_acc),
  em3 = mean(data_filtered$em3_acc),
  em4 = mean(data_filtered$em4_acc),
  em5 = mean(data_filtered$em5_acc)
)
```

```{r}
prop_acc_filtered2 <- data_filtered %>% 
  summarize(
  n = n(),
  
  em1_prop = mean(em1_acc),
  em1_sd = sd(em1_acc),
  em1_se = em1_sd/sqrt(n),
  em1_ci_lower = em1_prop - 1.96 * em1_se,
  em1_ci_upper = em1_prop + 1.96 * em1_se,
  
  em2_prop = mean(em2_acc),
  em2_sd = sd(em2_acc),
  em2_se = em2_sd/sqrt(n),
  em2_ci_lower = em2_prop - 1.96 * em2_se,
  em2_ci_upper = em2_prop + 1.96 * em2_se,
  
  em3_prop = mean(em3_acc),
  em3_sd = sd(em3_acc),
  em3_se = em3_sd/sqrt(n),
  em3_ci_lower = em3_prop - 1.96 * em3_se,
  em3_ci_upper = em3_prop + 1.96 * em3_se,
  
  em4_prop = mean(em4_acc),
  em4_sd = sd(em4_acc),
  em4_se = em4_sd/sqrt(n),
  em4_ci_lower = em4_prop - 1.96 * em4_se,
  em4_ci_upper = em4_prop + 1.96 * em4_se,
  
  em5_prop = mean(em5_acc),
  em5_sd = sd(em5_acc),
  em5_se = em5_sd/sqrt(n),
  em5_ci_lower = em5_prop - 1.96 * em5_se,
  em5_ci_upper = em5_prop + 1.96 * em5_se
  )

```

Validation trials were held out from the model training when estimating each embedding. We use these trials to estimate human agreement on trials and to provide a reference for each model's performance. Validation judgement were consistend across all participants.
```{r}
data_val <- data_all %>% 
  filter(type == "validation")

prop_acc_val <- tibble(
  em1 = mean(data_val$em1_acc),
  em2 = mean(data_val$em2_acc),
  em3 = mean(data_val$em3_acc),
  em4 = mean(data_val$em4_acc),
  em5 = mean(data_val$em5_acc)
)
  
```

```{r}
val <- data %>% 
  filter(type == "validation")
```

```{r}
majority_votes <- val %>%
  group_by(head, left, right) %>%
  count(winner, name = "n") %>%
  arrange(head, left, right, desc(n)) %>%
  slice(1) %>%
  ungroup() %>%
  rename(majority = winner)
```

```{r}
val <- val %>% 
  left_join(majority_votes %>% select(head, left, right, majority), by = c("head", "left", "right"))
```

```{r}
val <- val %>% 
  mutate(
    maj_agree = if_else(winner == majority, 1, 0),
    prop_agree = mean(maj_agree)
  )
```

```{r}
val_sum <- val %>% 
  group_by(head, left, right) %>% 
  summarise(
    prop_agree = mean(maj_agree)
  ) %>% 
  ungroup()
```

```{r}
# val_sum <- val %>% 
#   group_by(head, left, right) %>% 
#   summarise(
#     prop_agree = mean(maj_agree),
#     sd_prop_agree = sd(maj_agree),
#     n = n(),
#     se_prop_agree = sd_prop_agree/sqrt(n),
#     ci_lower = prop_agree - 1.96 * se_prop_agree,
#     ci_upper = prop_agree + 1.96 * se_prop_agree
#   ) %>% 
#   ungroup()
```

```{r}
prop_acc_filtered <- prop_acc_filtered %>% 
  mutate(
    human = unique(val$prop_agree)
  )
```

```{r}
# prop_acc_filtered2 <- prop_acc_filtered2 %>% 
#   mutate(
#     human = unique(val$prop_agree)
#   )
```

```{r}
# prop_long <- prop_acc_filtered %>%
#   pivot_longer(
#     cols = everything(),
#     names_to = "model",
#     values_to = "prop_agree"
#   )
```

```{r}
prop_long <- prop_acc_filtered2 %>%
  pivot_longer(
    cols = -n,
    names_to = c("model", "stat"),
    names_pattern = "(em[0-9]+)_(.*)",
    values_to = "value"
  )
```

```{r}
prop_long <- prop_long %>%
  pivot_wider(
    names_from = stat,
    values_from = value
  )
```

```{r}
human  <- prop_acc_filtered %>% select(human)
human <- human$human
```

## plot

```{r}
p_agree <- ggplot(prop_long,
       aes(x = model,
           y = prop)) +
  geom_hline(yintercept = human, linetype = "dashed") +
  annotate(
  "text",
  x = Inf,
  y = human,
  label = "Human agreement",
  vjust = 1.5,
  hjust = 1.1
)+
  geom_col(width = 0.5, fill = "grey") +
  geom_errorbar(
    aes(ymin = ci_lower,
        ymax = ci_upper),
    height = 0.2,
    width = .1,
    color = "black",
    linewidth = 0.2 
  ) +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+ 
  labs(
    title = "Model Predictions",
    x = "Model",
    y = "Proportion agree with human judgement"
  ) +
 theme_classic()

p_agree
```

```{r}
ggsave("../plots/embedding_pred.pdf", plot = p_agree, width = 3, height = 2)
```

# 5 Dimension Embedding

We do PCA on the highest dimension embedding to pair with out loss data to determine the final dimension solution to use for out space. 
```{r}
#remove labels, so just embedding matrix
em5_matrix <- em5 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result5 <- prcomp(em5_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores5 <- as.data.frame(pca_result5$x)
```

```{r}
# Add words back in
pca_scores5$word <- em5$word

```

```{r}
write.csv(pca_scores5, "pca_5_dim.csv", row.names = FALSE)
```

```{r}
# Eigenvalues

eigenvalues5 <- (pca_result5$sdev)^2

```

```{r}
pca_df5 <- data.frame(
  Principal_Component = 1:length(eigenvalues5),
  Eigenvalue = eigenvalues5
)

```

```{r}
# Scree plot

ggplot(pca_df5,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point(size = .5) +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance5 <- pca_result5$sdev^2 /
  sum(pca_result5$sdev^2)
cumulative_variance5 <- cumsum(explained_variance5)
explained_variance_df5 <- data.frame(
  Dimension = 1:length(explained_variance5),
  ExplainedVariance = explained_variance5,
  CumulativeVariance = cumulative_variance5
)

print(explained_variance_df5)

# Plot cumulative variance explained

p5_variance <- ggplot(explained_variance_df5,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             size = .5,
             color = "blue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p5_variance

```

```{r}
ggsave("../plots/PCA_reduction.pdf", plot = p5_variance, width = 3, height = 2)
```

# 4 Dimension Embedding
Based on the relative variance explained in each PC, the loss, and model agreement with human responses, we selected the 4 dimension solution. We do PCA on the 4D embedding to rotate the space for interpretable axes. 

```{r}
#remove labels, so just embedding matrix
em4_matrix <- em4 %>%
  select(starts_with("dim_"))
```

```{r}
#run PCA
pca_result4 <- prcomp(em4_matrix, center = TRUE, scale. = TRUE)

```

```{r}
# PCA scores
pca_scores4 <- as.data.frame(pca_result4$x)
```

```{r}
# Add words back in
pca_scores4$word <- em4$word

```

```{r}
write.csv(pca_scores4, "pca_4_dim.csv", row.names = FALSE)
```

```{r}
#switch signs of PC1 and PC3 so correlations from scales are positive
pca_scores4 <- pca_scores4 %>% 
  mutate(
    PC1 = PC1*(-1),
    PC3 = PC3*(-1)
  )

```

```{r}
# Eigenvalues

eigenvalues4 <- (pca_result4$sdev)^2

```

```{r}
pca_df4 <- data.frame(
  Principal_Component = 1:length(eigenvalues4),
  Eigenvalue = eigenvalues4
)

```

```{r}
# Scree plot

ggplot(pca_df4,
       aes(x = Principal_Component,
           y = Eigenvalue)) +
  geom_line() +
  geom_point() +
  labs(title = "Principal Components vs Eigenvalues",
       x = "Principal Component",
       y = "Eigenvalue") +
  theme_minimal()

```

```{r}
# Variance explained

explained_variance4 <- pca_result4$sdev^2 /
  sum(pca_result4$sdev^2)
cumulative_variance4 <- cumsum(explained_variance4)
explained_variance_df4 <- data.frame(
  Dimension = 1:length(explained_variance4),
  ExplainedVariance = explained_variance4,
  CumulativeVariance = cumulative_variance4
)

print(explained_variance_df4)

# Plot cumulative variance explained

p4_variance <- ggplot(explained_variance_df4,
       aes(x = Dimension)) +
  geom_line(aes(y = CumulativeVariance),
            color = "blue") +
  geom_point(aes(y = CumulativeVariance),
             color = "blue") +
  scale_y_continuous(limits = c(0, 1), expand = c(0,0))+
  labs(
    title = "Cumulative Variance Explained by PCA",
    x = "Number of Dimensions",
    y = "Cumulative Variance Explained"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_variance

```

```{r}
pca_scores4 <- pca_scores4 %>%
  left_join(colors, by = "word")
```

```{r}
# PC1 vs PC2

p4_pc12 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC2,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  # scale_x_continuous(
  # limits = c(min(pca_scores4$PC1), max(pca_scores4$PC1) + 0.3)
  # ) +
  scale_y_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  labs(
    title = "PCA Space: PC1 vs PC2",
    x = "PC1",
    y = "PC2"
  ) +
theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc12
```

```{r}
# PC1 vs PC3

p4_pc13 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC3,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_y_continuous(
  limits = c(min(pca_scores4$PC3), max(pca_scores4$PC3) +.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC1 vs PC3",
    x = "PC1",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc13
```

```{r}
# PC1 vs PC4

p4_pc14 <- ggplot(pca_scores4,
       aes(x = PC1,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC1 vs PC4",
    x = "PC1",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc14
```

```{r}
# PC2 vs PC3

p4_pc23 <- ggplot(pca_scores4,
       aes(x = PC2,
           y = PC3,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_x_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC2 vs PC3",
    x = "PC2",
    y = "PC3"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc23
```

```{r}
# PC2 vs PC4

p4_pc24 <- ggplot(pca_scores4,
       aes(x = PC2,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  scale_x_continuous(
  limits = c(min(pca_scores4$PC2), max(pca_scores4$PC2) + 0.3)
  ) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC2 vs PC4",
    x = "PC2",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc24
```

```{r}
# PC3 vs PC4

p4_pc34 <- ggplot(pca_scores4,
       aes(x = PC3,
           y = PC4,
           label = word,
           color = color)) +
  geom_vline(xintercept = 0, color = "grey", linewidth = .3) +
  geom_hline(yintercept = 0, color = "grey", linewidth = .3) +
  geom_point(size = 1) +
  geom_text(vjust = -0.5, size = 3) +
  scale_color_identity() +
  # geom_hline(yintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  # geom_vline(xintercept = 0,
  #            linetype = "dashed",
  #            color = "grey50") +
  labs(
    title = "PCA Space: PC3 vs PC4",
    x = "PC3",
    y = "PC4"
  ) +
 theme_classic() +
  theme(
    panel.border = element_rect(
      color = "black",
      fill = NA,
      linewidth = 1
    )
  )

p4_pc34
```

```{r}
ggsave("../plots/dim4_pca_pc1_pc2.png", plot = p4_pc12, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc3.png", plot = p4_pc13, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc4.png", plot = p4_pc14, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc2_pc3.png", plot = p4_pc23, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc2_pc4.png", plot = p4_pc24, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc3_pc4.png", plot = p4_pc34, width = 10, height = 10, dpi = 300)

ggsave("../plots/dim4_pca_pc1_pc2.pdf", plot = p4_pc12, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc1_pc3.pdf", plot = p4_pc13, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc1_pc4.pdf", plot = p4_pc14, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc2_pc3.pdf", plot = p4_pc23, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc2_pc4.pdf", plot = p4_pc24, width = 3.5, height = 3.5)

ggsave("../plots/dim4_pca_pc3_pc4.pdf", plot = p4_pc34, width = 3.5, height = 3.5)
```

#Participants and Timing
During piloting we checked approximately how long the experiment took each participant to share in our recruitment message.

```{r}
subj <- read_csv("../data/participants.csv")
```

```{r}
timing <- subj
```

```{r}
# Convert to datetime and calculate duration

timing <- timing %>%
  mutate(
    start_time = ymd_hms(started_at),
    stop_time  = ymd_hms(completed_at),
    
    # duration in seconds
    duration_sec = as.numeric(difftime(completed_at, started_at, units = "secs")),
    
    # duration in minutes
    duration_min = as.numeric(difftime(completed_at, started_at, units = "mins")),
    

    # duration in hours
    duration_hr = as.numeric(difftime(completed_at, started_at, units = "hours"))
  )

print(timing)

```

```{r}
timing <- timing %>%
  mutate(
    duration = completed_at - started_at
  )

print(timing$duration)
```

```{r}
mean(as.numeric(timing$duration_min)[as.numeric(timing$duration_min) < 200],
     na.rm = TRUE)
```
